---
title: "HW5"
author: "Ian Driscoll"
date: "9/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 5A.
```{r}
# X bar
x = seq(-10, 10, length=1000)
y = dnorm(x, mean=0, sd=2)
plot(x, y, type="l", lwd=1, main = "Sampling Distribution of X bar")
# Sigma hat
curve(4 * dchisq(x,df = 24),xlim=c(0,60),ylim=c(0,0.6),ylab="Chi Square Density", main = "Distribution of Sigma Hat")
```

## 5D.
```{r}
data = c(5.3299, 4.2537, 3.1502, 3.7032, 1.6070, 6.3923, 3.1181, 6.5941, 3.5281, 4.7433, 0.1077, 1.5977, 5.4920, 1.7220, 4.1547, 2.2799)
moment1 = mean(data)
sample_var = var(data)
```

## 5G.
```{r}
library(dplyr)
library(ggplot2)
scores = read.table("Desktop/data.scores.txt",header=TRUE)
scores = filter(scores, f > 0 & m > 0)
scores = data.frame(m = scores$m * 2, f = scores$f)

# Part A.

plot1 = ggplot(scores) + geom_histogram(aes(x = m, fill = 'midterm', alpha = 0.2)) + geom_histogram(aes(x = f, fill = 'final', alpha = 0.2))
plot1
```

```{r}
# Part B.

boxplot(x = scores$m, scores$f, xlab = "Midterm vs. Final", ylab = "Score (Out of 40)")                                         
```

The spread is very similar for both the midterm and final, with people on average doing slightly better on the final. The upper third quartile is a smaller range for the final than the midterm, and the lower second quartile is larger than on the midterm. Both are centered around 24-26.

## 5H

```{r}
# Part A.

random500 = rnorm(500, 10, 3)
qqnorm(random500)
```
My plot does appear to be linear.

```{r}
# Part B.
qqnorm(scores$f)
```
The histogram does not appear normal just like this plot, they both appear skewed to the left.

## 5I
```{r}
stem(scores$f)
stem(scores$f, scale = 0.5)
stem(scores$f, scale = 2)
```
The plot with scale 2 is the most detailed in the shape of the distribution of scores.

## 5J

```{r}
# Part A.
ggplot(scores) + geom_point(aes(x = m, y = f)) + geom_abline(slope = 1, intercept = 0)

```
Eyeballing the graph, I would say the percentage who gained from the midterm is greater than 50%.

```{r}
# Part B.
change = scores$f - scores$m
percent = length(change[change > 0]) / length(scores$f) * 100
percent
```
##### Part C.

For this plot, I would do a slightly less steep line with a larger y-intercept, since the line would have to be shifted up due to final scores being higher on average overall, and it would be less steep than a slope of 1 since scores would have to level off as they got closer and closer to 40 since they are starting from a higher y coordinate.




