---
title: "HW10-R"
author: "Ian Driscoll"
date: "11/20/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 10F.

### a.

```{r}
bodytemp = read.delim("bodytemp.txt", sep = " ")
males = bodytemp[bodytemp$gender == 1, ]
females = bodytemp[bodytemp$gender == 2, ]
plot(males$temperature, males$rate)
plot(females$temperature, females$rate)
```

It seems that there is a slight positive correlation between body temperature and heart rate.

### b.

```{r}
library(ggplot2)
ggplot(bodytemp) + geom_point(aes(x = temperature, y = rate, col = gender))
```

It seems that there is a slightly more positive correlation among females.

### c.

```{r}
male_lm = lm(formula = rate ~ temperature, data = males)
summary(male_lm)
plot(males$temperature, male_lm$residuals)
```

There appears to be no linear correlation in the residual plot, which shows that the simple linear regression is a good fit. From the summary of the regression, the slope is 1.645 and the standard error is 1.039.

### d.

```{r}
female_lm = lm(formula = rate ~ temperature, data = females)
summary(female_lm)
plot(females$temperature, female_lm$residuals)
```

There appears to be no linear correlation in the residual plot, which shows that the simple linear regression is a good fit. From the summary of the regression, the slope is 3.128 and the standard error is 1.316.




