---
title: "HW 02 - Shot Charts"
author: "Ian Driscoll"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```

## 5. Effective Percentages
```{r}
library(dplyr)
library(ggplot2)
dat = read.csv('../data/shots-data.csv')

# table of players and total shot attempts

dat %>%
  group_by(name) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

# table with player, total shot attempts, made shot, and percentage 

iggy = nrow(filter(dat, name == 'Andre Iguodala' & shot_made_flag == 'made shot'))
green = nrow(filter(dat, name == 'Draymond Green' & shot_made_flag == 'made shot'))
durant = nrow(filter(dat, name == 'Kevin Durant' & shot_made_flag == 'made shot'))
thompson = nrow(filter(dat, name == 'Klay Thompson' & shot_made_flag == 'made shot'))
curry = nrow(filter(dat, name == 'Stephen Curry' & shot_made_flag == 'made shot'))
made_shots = c(iggy, green, durant, thompson, curry)

dat %>%
  group_by(name) %>%
  summarise(total = n()) %>%
  mutate(made = made_shots) %>%
  mutate(perc_made = made / total) %>%
  arrange(desc(perc_made))

# 2 point shooting percentage

iggy2 = nrow(filter(dat, name == 'Andre Iguodala' & shot_made_flag == 'made shot' & shot_type == '2PT Field Goal'))
green2 = nrow(filter(dat, name == 'Draymond Green' & shot_made_flag == 'made shot' & shot_type == '2PT Field Goal'))
durant2 = nrow(filter(dat, name == 'Kevin Durant' & shot_made_flag == 'made shot' & shot_type == '2PT Field Goal'))
thompson2 = nrow(filter(dat, name == 'Klay Thompson' & shot_made_flag == 'made shot' & shot_type == '2PT Field Goal'))
curry2 = nrow(filter(dat, name == 'Stephen Curry' & shot_made_flag == 'made shot' & shot_type == '2PT Field Goal'))
made_shots2 = c(iggy2, green2, durant2, thompson2, curry2)

dat %>%
  group_by(name) %>%
  filter(shot_type == '2PT Field Goal') %>%
  summarise(total = n()) %>%
  mutate(made = made_shots2) %>%
  mutate(perc_made = made / total) %>%
  arrange(desc(perc_made))

# 3 point shooting percentage

iggy3 = nrow(filter(dat, name == 'Andre Iguodala' & shot_made_flag == 'made shot' & shot_type == '3PT Field Goal'))
green3 = nrow(filter(dat, name == 'Draymond Green' & shot_made_flag == 'made shot' & shot_type == '3PT Field Goal'))
durant3 = nrow(filter(dat, name == 'Kevin Durant' & shot_made_flag == 'made shot' & shot_type == '3PT Field Goal'))
thompson3 = nrow(filter(dat, name == 'Klay Thompson' & shot_made_flag == 'made shot' & shot_type == '3PT Field Goal'))
curry3 = nrow(filter(dat, name == 'Stephen Curry' & shot_made_flag == 'made shot' & shot_type == '3PT Field Goal'))
made_shots3 = c(iggy3, green3, durant3, thompson3, curry3)

dat %>%
  group_by(name) %>%
  filter(shot_type == '3PT Field Goal') %>%
  summarise(total = n()) %>%
  mutate(made = made_shots3) %>%
  mutate(perc_made = made / total) %>%
  arrange(desc(perc_made))
```

# 6. Shooting Distances
```{r}
a = dat %>%
  group_by(shot_distance) %>%
  summarise(made_shot_prop = length(shot_made_flag[shot_made_flag=="made shot"]) / n())
  
ggplot(data = a) + geom_point(aes(x = shot_distance, y = made_shot_prop))
```

From the plot we can observe that as distance increased, proportion does go down for the most part. It appears that beyond 32-34 feet the chance of making a shot drops to practically zero. Distances within 3 feet, and a few distances between 16 and 22 feet have over a 50 percent chance of a shot being made.

# 7. Total Shots by Minute of Occurrence
```{r}
player_distances = dat %>%
  group_by(name, minute) %>%
  summarise(total_shots = n())

ggplot(data = player_distances) + 
  geom_rect(aes(xmin = 1, ymin = 0, xmax = 12, ymax = Inf), fill = 'grey97') + 
  geom_rect(aes(xmin = 24, ymin = 0, xmax = 36, ymax = Inf), fill = 'grey97') + 
  geom_point(aes(x = minute, y = total_shots), color = 'steelblue2', size = 0.8) + 
  facet_wrap(~ name) + 
  geom_path(aes(x = minute, y = total_shots), color = 'steelblue2', size = 0.4) + 
  theme_minimal() + 
  scale_x_continuous(breaks = c(1, 12, 24, 36, 48)) + 
  ggtitle('Total number of shots (by minute of occurrence)') + 
  xlab('minute') + 
  ylab('total number of shots')

```


