---
title: "Warm-Up 2 Ian Driscoll"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
```{r}
load("nba2017-salary-points.RData")
```

### Salary Miscellaneous Information

```{r}
salary_millions = salary / 1000000
pos_fac = as.factor(position)
levels(pos_fac) = c("center", "power_fwd", "point_guard", "small_fwd", "shoot_guard")
summary(pos_fac)
summary(salary)
mean(salary)
sd(salary)
min(salary)
max(salary)
quantile(salary)
```
  
The typical values for salary are centered around 6.2 million, and it is skewed to the right. For salary, there are a few outliers above 20 million. The histogram and boxplot are displayed below.  

```{r}
hist(salary)
boxplot(salary)
```

### Team Information 

```{r}
table(team)
freq = table(team) / length(team)
freq
barplot(freq)
```
Each team has either 14 or 15 players, and the distribution is uniform, with each relative frequency around 3%.  

### Scatterplot of Points and Salar

```{r}
plot.new()
plot.window(c(0, 2600), c(0, 30))
axis(1, labels = TRUE)
axis(2, labels = TRUE)
title(main = "Points vs. Salary", xlab = "Points", ylab = "Salary (millions)")
points(points, salary_millions)
```

### Finding Correlation Between Points and Salary

```{r}
# First we must find the mean of each vector.
mean_points = sum(points) / length(points)
mean_salary = sum(salary_millions) / length(salary_millions)
# Next comes calculating the standard deviations.
sd_salary = sqrt((sum((salary_millions - mean_salary) ^ 2)) / (length(salary_millions) - 1))
sd_points = sqrt((sum((points - mean_points) ^ 2)) / (length(points) - 1))
# We also must find covariance.
cov_pts_salary = sum((points - mean_points) * (salary_millions - mean_salary)) / (length(points) - 1)
#Plugging all of our values into our correlation formula gives us our answer.
cor_pts_salary = cov_pts_salary / (sd_points * sd_salary)
cor_pts_salary
```

### Comments and Reflections

- Formatting Markdown was stil pretty tedious, the syntax will just take some getting used to.
- Assigning values and coming to the correlation manually was pretty simple for me.
- I did not need any help for this assignment other than quick google searches for R commands.
- This assignment took me about 2 hours in total.
- The most time consuming part was formatting the Markdown file.
- I understood everything pretty well.
- Nothing was super frustrating in particular, other than my lack of comfort with Markdown syntax.