---
title: "Dieroller"
author: "Ian Driscoll"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to dieroller}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(dieroller)
```

## Coin Tossing

The package `"dieroller"` is a use for rolling a 6-sided die and plotting the results.


## Creating a `die`

The first step is to create a `"die"` object with the function `die()`:

```{r}
testdie = die()
testdie
```

By default, `die()` creates a _fair_ die with standard sides 1-6, and with probability values for each side of 1/6.

You can change the defaults with arguments `sides` and `prob`:

```{r}
loaded <- die(sides = c('x', 'y', 'z', 'a', 'b', 'c'), prob = c(0.4, 0.2, 0.1, 0.1, 0.1, 0.1))
loaded
```


## Rolling a die

Once you have defined a `"die"` object, you can use the function `roll()` to
_roll_ a die a given number of `times`. The output will be an object of class 
`"roll"`, which will contain the vector of `rolls`:

```{r}
roll100 <- roll(testdie, times = 100)
roll100
```

The `summary()` function gives you a nice table of frequencies:
```{r}
summary(roll100)
```


## Plot of relative frequencies

You can use the function `plot()` to create a graohic with the relative 
frequencies of a series of die rolls:

```{r, fig.show='hold'}
plot(roll100)
```
